Probabilistic Catalogs from Crowded Star Fields

### 2012-07-15 paper/ms.tex
Some more writing

A fundamental problem in astronomy is the construction of {\it catalogs} from
raw image data. Essentially, the question being answered is ``given the image
we have obtained, what objects are present in the field and what are their
properties?''. Standard tools include \texttt{SExtractor} \citep{sextractor}.

\section{Probabilistic Catalogs}

\citet{2011MNRAS.415.3462F} have attempted this, by replacing the complex
$N$-component model with a single-component model, and using the
single-component
results to reconstruct what the $N$-component results would have been. However,
this will not work well in situations where there is significant confusion
(i.e. two or more sources overlap) and also does not allow for unknown $N$,
which is the general situation. What we really need is a variable dimension
sampler such as reversible jump MCMC \citep{rjmcmc}.

\begin{itemize}
\item Catalogs as models for what's out there \\
\item Standard techniques for constructing catalogs, and under what
cicrumstances they stop working \\
\item 
\end{itemize}

\section{Bayesian Inference}
We use Bayesian Inference. Suppose there exist unknown parameters (denoted
collectively by $\theta$) and we expect to obtain some data $D$. Our prior
knowledge (or uncertainty) about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta)
\end{equation}
We also model how the parameters give rise to the data, via a {\it sampling
distribution}:
\begin{equation}
p(D|\theta)
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions, one for each possible value of $\theta$. Note that
the sampling distribution is also an assertion of a prior state of knowledge.
In this case it models our prior information about the fact that the parameters
$\theta$ are going to influence the data in some way (otherwise this data
would be irrelevant).

@@

@@

>>>>

A fundamental problem in astronomy is the construction of {\it catalogs} from
raw image data. {\bf The catalogs are then used for...}
Standard tools for generating catalogs include \texttt{SExtractor}
\citep{sextractor}, {\bf more refs}.

However, standard methods for constructing catalogs can have difficulty in some
challenging situations. For example, when multiple sources overlap partially
or completely, it can be difficult to determine how many sources are present,
and how much flux belongs to each source. In principle, uncertainty should be
taken into account -- instead of simply estimating the position and flux of each
object, we should fully describe the fact that sometimes we are not certain of
the position and flux of each source.

Essentially, the creation of a catalog is an attempt to answer the question
``given the image we have obtained, what objects are present in the field and
what are their properties?''. This motivates a probabilistic (Bayesian)
approach to making catalogs.

\citet{2011MNRAS.415.3462F} have attempted this, by replacing the complex
$N$-component model with a single-component model, and using the
single-component
results to reconstruct what the $N$-component results would have been. However,
this will not work well in situations where there is significant confusion
(i.e. two or more sources overlap) and also does not allow for unknown $N$,
which is the general situation. What we really need is a variable dimension
model, where $N$ is an unknown quantity to be inferred from the data.
The computational implementation of these models will require tools such as reversible jump Markov Chain Monte Carlo \citep{rjmcmc}.

\section{Bayesian Inference}
To quantitatively model uncertainties, Bayesian Inference is the appropriate
framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters (denoted
collectively by $\theta$) and we expect to obtain some data $D$. Our prior
knowledge (or uncertainty) about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta)
\end{equation}
We also model how the parameters give rise to the data, via a {\it sampling
distribution}:
\begin{equation}
p(D|\theta)
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions, one for each possible value of $\theta$. Note that
the sampling distribution is also an assertion of a prior state of knowledge.
In this case it models our prior information about the fact that the parameters
$\theta$ are going to influence the data in some way (otherwise this data
would be irrelevant).

@@

\bibitem[Caticha(2009)]{caticha} Caticha, A.\ 2009.\ 
Quantifying Rational Belief.\ American Institute of Physics Conference 
Series 1193, 60-68.

\bibitem[Cox(1946)]{cox} Cox, R.~T., 1946, Probability, Frequency, and Reasonable Expectation. 1946. American Journal of Physics 14 14, 1-13.

@@

\bibitem[Jaynes(2003)]{jaynes} Jaynes, E.~T., 2003, Probability Theory: The
Logic of Science, ISBN 0521592712, Cambridge University Press, June 2003.
